Höglinger et al. 2017

### 2025-02-25 IndividualStudies/DH_2017.qmd
Found a silly mistake in my Volcano limits function
Adding GO analysis to DH_2017

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(pheatmap)
library(here)
library(heatmaply)
library(ggvenn)
library(scales)

@@

DH_2017_data <- read.csv(paste0(here(),"/IndividualStudies/DataTables/DH_PNAS_2017_download.csv")) |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>
  arrange(gene_name) |>
  group_by(gene_name) |>
  summarise(PSM_DAG = mean(PSM_DAG),
            PSM_FA = mean(PSM_FA),
            PSM_Sph = mean(PSM_Sph))
```

@@

#### Heat map
:::{.border}
```{r, echo=FALSE, fig.height=10, warning=FALSE}

DH_2017_data_long <- DH_2017_data |>
  pivot_longer(cols = starts_with("PSM_"),
              names_to = "variable",
              names_prefix = "PSM_",
              values_to = "value") |>
  mutate(name = gene_name) |>
  select(name, variable, value)
  # mutate(OriginalValues = value) |>
  # mutate(value = case_when(value < 20 ~ value,
  #                           value >= 20 ~ 20))

DH_2017_data_long$variable <- factor(pull(DH_2017_data_long, variable), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

# Create the heatmap
heatmaply(long_data = DH_2017_data_long,
          xlab = "Probe",
          ylab = "Gene name",
          label_names = c("Gene name", "Probe", "PSM count"),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradientn(colours = c("grey95", "purple", "orange", "orange", "orange", "orange", "orange", "red", "red","red"), breaks = c(0, 5, 20, 200)),
          showticklabels = F)

@@

:::{.border}
```{r, echo=FALSE, warning=FALSE, width = 8}

# Ensure the data is in the correct format
DH_2017_data_AddingRownames <- as.data.frame(DH_2017_data)
rownames(DH_2017_data_AddingRownames) <- DH_2017_data_AddingRownames[,1]
DH_2017_data_AddingRownames[,1] <- NULL

Sph_proteins <- DH_2017_data_AddingRownames |>
  filter(PSM_Sph > 0) |>
  select(PSM_Sph)

FA_proteins <- DH_2017_data_AddingRownames |>
  filter(PSM_FA > 0) |>
  select(PSM_FA)

DAG_proteins <- DH_2017_data_AddingRownames |>
  filter(PSM_DAG > 0) |>
  select(PSM_DAG)

combined_venn_proteins <- list("Sph proteins" = rownames(Sph_proteins), "FA proteins" = rownames(FA_proteins), "DAG proteins" = rownames(DAG_proteins))

ggvenn <- ggvenn(combined_venn_proteins,
                      show_percentage = FALSE,
                      show_elements = FALSE,
                      fill_color = c("orange", "purple", "blue")) +
        labs(x = "", y = "") +
        theme_minimal()

@@

```{r, echo=FALSE, warning=FALSE}

DH_2017_data_reactive <- crosstalk::SharedData$new(DH_2017_data)

@@

## Displays the reactive dataframe
datatable(DH_2017_data_reactive, options = list(pageLength = 20))

>>>>

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(here)
library(heatmaply)
library(ggvenn)
library(scales)

@@

###### Wrangling the raw dataset ######
DH_2017_data <- read.csv(paste0(here(),"/IndividualStudies/DataTables/DH_PNAS_2017_download.csv")) |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  dplyr::select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>
  arrange(gene_name) |>
  group_by(gene_name) |>
  summarise(PSM_DAG = mean(PSM_DAG),
            PSM_FA = mean(PSM_FA),
            PSM_Sph = mean(PSM_Sph))

###### Wrangling for the DataTable and the GO ######

###### Wrangling for the HeatMap ######

DH_2017_data_HeatMap <- DH_2017_data |>
  pivot_longer(cols = starts_with("PSM_"),
              names_to = "variable",
              names_prefix = "PSM_",
              values_to = "value") |>
  mutate(name = gene_name) |>
  dplyr::select(name, variable, value)

DH_2017_data_HeatMap$variable <- factor(pull(DH_2017_data_HeatMap, variable), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

###### Wrangling for the Venn Diagram ######

# Separates out each of the genesets, then forces them back together in a vectorized list.
Sph_proteins <- DH_2017_data |>
  filter(PSM_Sph > 0) |>
  dplyr::select(gene_name)

FA_proteins <- DH_2017_data |>
  filter(PSM_FA > 0) |>
  dplyr::select(gene_name)

DAG_proteins <- DH_2017_data |>
  filter(PSM_DAG > 0) |>
  dplyr::select(gene_name)

DH_2017_VennData <- list("Sph proteins" = as.list(Sph_proteins$gene_name), "FA proteins" = as.list(FA_proteins$gene_name), "DAG proteins" = as.list(DAG_proteins$gene_name))

@@

#### Heat map

Depicted below are the average PSM count for two replicates each of UV-irradiated samples treated with trifunctionalized Sph, DAG, and FA.

:::{.border}

```{r Making Heat Map, echo=FALSE, fig.height=10, warning=FALSE}

# Create the heatmap
heatmaply(long_data = DH_2017_data_HeatMap,
          xlab = "Probe",
          ylab = "Gene name",
          label_names = c("Gene name", "Probe", "PSM count"),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradientn(colours = c("grey95", "purple", "orange", "orange", "orange", "orange", "orange", "red", "red","red"), breaks = c(0, 5, 20, 200)),
          showticklabels = F)

@@

:::{.border}

```{r Making Venn Diagram, echo=FALSE, warning=FALSE, width = 8}

ggvenn <- ggvenn(DH_2017_VennData,
                      show_percentage = FALSE,
                      show_elements = FALSE,
                      fill_color = c("orange", "purple", "blue")) +
        labs(x = "", y = "") +
        theme_minimal()

@@

## Gene Ontology Analysis

```{r GO data wrangling, echo=FALSE, warning=FALSE}

# ID_LUT <- clusterProfiler::bitr(geneID= DH_2017_data$gene_name, 
#                                 fromType = "SYMBOL",
#                                 toType = "ENTREZID",
#                                 OrgDb = org.Hs.eg.db
# )

# names(ID_LUT) <- c("gene_name", "ENTREZID")

# # Merging ENTREZIDs to data
# DH_2017_data_GO <- left_join(DH_2017_data, ID_LUT)

# ego_results_Identification_CC <- NULL

# # Attempts to identify pathways in the dataset
# try(
#   ego_results_Identification_CC <- clusterProfiler::compareCluster(ENTREZID ~ VennRegion,
#                                     data = DH_2017_data_GO, fun = "enrichGO",
#                                     OrgDb = org.Hs.eg.db,
#                                     keyType = "ENTREZID",
#                                     ont = "CC",
#                                     readable = TRUE,
#                                     universe = ID_LUT$ENTREZID)
#                                     )

# # If no pathway identified, returns text warning user.
# if (is.null(ego_results_Identification_CC)) {
#   print("No pathways identified among enriched hits and candidates.")
# }

# # If pathways identified, produces both plots for viewing.
# if (!is.null(ego_results_Identification_CC))
#   {
#   ego_results_Identification_CC_table <- ego_results_Identification_CC@compareClusterResult
#   ego_results_Identification_CC_table <- ego_results_Identification_CC_table %>% 
#     group_by(ID, VennRegion) %>%
#     mutate(odds_ratio = calculateFE(GeneRatio, BgRatio))
#   ego_results_Identification_CC_table$Description <- factor(ego_results_Identification_CC_table$Description, levels = unique(rev(ego_results_Identification_CC_table$Description)))

#   ego_sub <- ego_results_Identification_CC_table %>% 
#     group_by(VennRegion) %>%
#     slice_head(n = 10)

#   # Makes the dot plot
#   dot <- ggplot(data = ego_sub, aes(VennRegion, Description)) +
#     geom_point(aes(size = odds_ratio, colour = -log10(p.adjust))) +
#     theme_bw(base_size = 12) +
#     scale_colour_gradientn(colours = c("#377eb8", "#984ea3", "#e41a1c", "#ff7f00", "#ffff33"), ) +
#     # ggtitle("Cellular compartment") +
#     ylab("") +
#     xlab("Enrichment Profile") +
#     theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

#   dot <- ggplotly(dot)
#   # Save as HTML (adjust the path as needed)
#   saveWidget(dot, file = paste0(here(), "/IndividualStudies/GO_plots/DH_2017_CC_DOTPLOT.html"), selfcontained = TRUE)

#   try(
# 			# Makes the cnet plot
#     cnet <- clusterProfiler::cnetplot(ego_results_Identification_CC, categorySize = "pvalue") +
#       # ggtitle("Cellular compartment") +
#         customPlot
#     )

#   if(!is.null(cnet)){
#       ggsave(paste0(here(), "/IndividualStudies/GO_plots/DH_2017_CC_CNET.png"),
#            width = 12, height = 12)
#   } else {
#     print("Cnet plot failed in production.")
#   }
# }

# ego_results_Identification_MF <- NULL

# # Attempts to identify pathways in the dataset
# try(
#   ego_results_Identification_MF <- clusterProfiler::compareCluster(ENTREZID ~ VennRegion,
#                                     data = DH_2017_data_GO, fun = "enrichGO",
#                                     OrgDb = org.Hs.eg.db,
#                                     keyType = "ENTREZID",
#                                     ont = "MF",
#                                     readable = TRUE,
#                                     universe = ID_LUT$ENTREZID)
#                                     )

# # If no pathway identified, returns text warning user.
# if (is.null(ego_results_Identification_MF)) {
#   print("No pathways identified among enriched hits and candidates.")
# }

# # If pathways identified, produces both plots for viewing.
# if (!is.null(ego_results_Identification_MF))
#   {
  # ego_results_Identification_MF_table <- ego_results_Identification_MF@compareClusterResult
  # ego_results_Identification_MF_table <- ego_results_Identification_MF_table %>% 
  #   group_by(ID, VennRegion) %>%
  #   mutate(odds_ratio = calculateFE(GeneRatio, BgRatio))
  # ego_results_Identification_MF_table$Description <- factor(ego_results_Identification_MF_table$Description, levels = unique(rev(ego_results_Identification_MF_table$Description)))

  # ego_sub <- ego_results_Identification_MF_table %>% 
  #   group_by(VennRegion) %>%
  #   slice_head(n = 10)

  # # Makes the dot plot
  # dot <- ggplot(data = ego_sub, aes(VennRegion, Description)) +
  #   geom_point(aes(size = odds_ratio, colour = -log10(p.adjust))) +
  #   theme_bw(base_size = 12) +
  #   scale_colour_gradientn(colours = c("#377eb8", "#984ea3", "#e41a1c", "#ff7f00", "#ffff33"), ) +
  #   # ggtitle("Cellular compartment") +
  #   ylab("") +
  #   xlab("Enrichment Profile") +
  #   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

  # dot <- ggplotly(dot)
  # # Save as HTML (adjust the path as needed)
  # saveWidget(dot, file = paste0(here(), "/IndividualStudies/GO_plots/DH_2017_MF_DOTPLOT.html"), selfcontained = TRUE)

#   try(
# 			# Makes the cnet plot
#     cnet <- clusterProfiler::cnetplot(ego_results_Identification_MF, categorySize = "pvalue") +
#       # ggtitle("Cellular compartment") +
#         customPlot
#     )

#   if(!is.null(cnet)){
#       ggsave(paste0(here(), "/IndividualStudies/GO_plots/DH_2017_MF_CNET.png"),
#            width = 12, height = 12)
#   } else {
#     print("Cnet plot failed in production.")
#   }
# }

# ego_results_Identification_BP <- NULL

# # Attempts to identify pathways in the dataset
# try(
#   ego_results_Identification_BP <- clusterProfiler::compareCluster(ENTREZID ~ VennRegion,
#                                     data = DH_2017_data_GO, fun = "enrichGO",
#                                     OrgDb = org.Hs.eg.db,
#                                     keyType = "ENTREZID",
#                                     ont = "BP",
#                                     readable = TRUE,
#                                     universe = ID_LUT$ENTREZID)
#                                     )

# # If no pathway identified, returns text warning user.
# if (is.null(ego_results_Identification_BP)) {
#   print("No pathways identified among enriched hits and candidates.")
# }

# If pathways identified, produces both plots for viewing.
# if (!is.null(ego_results_Identification_BP))
#   {
#   ego_results_Identification_BP_table <- ego_results_Identification_BP@compareClusterResult
#   ego_results_Identification_BP_table <- ego_results_Identification_BP_table %>% 
#     group_by(ID, VennRegion) %>%
#     mutate(odds_ratio = calculateFE(GeneRatio, BgRatio))
#   ego_results_Identification_BP_table$Description <- factor(ego_results_Identification_BP_table$Description, levels = unique(rev(ego_results_Identification_BP_table$Description)))

#   ego_sub <- ego_results_Identification_BP_table %>% 
#     group_by(VennRegion) %>%
#     slice_head(n = 10)

#   # Makes the dot plot
#   dot <- ggplot(data = ego_sub, aes(VennRegion, Description)) +
#     geom_point(aes(size = odds_ratio, colour = -log10(p.adjust))) +
#     theme_bw(base_size = 12) +
#     scale_colour_gradientn(colours = c("#377eb8", "#984ea3", "#e41a1c", "#ff7f00", "#ffff33"), ) +
#     # ggtitle("Cellular compartment") +
#     ylab("") +
#     xlab("Enrichment Profile") +
#     theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

#   dot <- ggplotly(dot)
#   # Save as HTML (adjust the path as needed)
#   saveWidget(dot, file = paste0(here(), "/IndividualStudies/GO_plots/DH_2017_BP_DOTPLOT.html"), selfcontained = TRUE)

#   try(
# 			# Makes the cnet plot
#     cnet <- clusterProfiler::cnetplot(ego_results_Identification_BP, categorySize = "pvalue") +
#       # ggtitle("Cellular compartment") +
#         customPlot
#     )

#   if(!is.null(cnet)){
#       ggsave(paste0(here(), "/IndividualStudies/GO_plots/DH_2017_BP_CNET.png"),
#            width = 12, height = 12)
#   } else {
#     print("Cnet plot failed in production.")
#   }
# }
```

**In beta: GO analysis still under development**

:::: {.panel-tabset}

#### Cell Compartment
:::{.border}
<iframe src="/IndividualStudies/GO_plots/DH_2017_CC_DOTPLOT.html" width="100%" height="500px"></iframe>

#### Molecular Function
:::{.border}
<iframe src="/IndividualStudies/GO_plots/DH_2017_MF_DOTPLOT.html" width="100%" height="500px"></iframe>

#### Biological Process
:::{.border}
<iframe src="/IndividualStudies/GO_plots/DH_2017_BP_DOTPLOT.html" width="100%" height="500px"></iframe>

:::
::::

:::{.border}
```{r Building Datatable, echo=FALSE, warning=FALSE}
DH_2017_data_reactive <- crosstalk::SharedData$new(DH_2017_data)

@@

## Displays the reactive dataframe
datatable(DH_2017_data_reactive, options = list(pageLength = 20))
```
